Program Transformations for miniKanren

### 2020-07-06 papers/aplas/text/evaluation.tex

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.6\textwidth}
    \begin{lstlisting}[label={doubleApp}, caption={Concatenation of three lists}, captionpos=b, frame=tb]
  let doubleAppend$^o$ xs ys zs res =
    fresh (ts) (
      append$^o$ xs ys ts /\ append$^o$ ts zs res)

  let rec append$^o$ xs ys zs = conde [
    (xs === [] /\ ys === zs);
    fresh (h t r) (
      xs === (h % t) /\
      zs === (h % r) /\
      append$^o$ t ys r)]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.6\textwidth}
    \begin{lstlisting}[label={doubleApp:ideal}, caption={Ideal implementation of concatenation of three lists}, captionpos=b, frame=tb]
  let doubleAppend$^o$ xs ys zs res = conde [
    (xs === [] /\ append$^o$ ys zs res);
    fresh (h t r) (
      xs === h % t /\
      res === h % r /\
      doubleAppend$^o$ t ys zs r)]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

\begin{table}
  \centering
  \begin{tabular}{c||c||c}
                  & forward & backward \\
  \hline\hline
  Original        & 34.8ms & 1.6ms \\ \hline
  \ecce           & 26.7ms & 1.7ms \\ \hline
  Ideal           & 26.9ms & 1.7ms \\ \hline
  ConsPD         & 47.6ms & 1.7ms
  \end{tabular}
  \caption{Evaluation results for doubleAppendo}
  \label{tbl:doubleApp}
\end{table}

\includegraphics{../data/doubleAppendo/da700log.pdf}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.85\textwidth}
\begin{lstlisting}[label={cpd:maxandlength}, caption={Maximum element and length of the list}, captionpos=b, frame=tb]
  let maxLength$^o$ xs m l = max$^o$ xs m /\ length$^o$ xs l
  let rec length$^o$ xs l =
    fresh (h t m) ( conde [
      (xs === [] /\ l === zero);
      (xs === h % t /\ l === succ m /\ length$^o$ t m)])

@@

  let rec le$^o$ x y b =
    fresh (x$_1$ y$_1$) ( conde [
      (x === zero /\ b === ^true);
      (x === succ x$_1$ /\ y === zero /\ b === ^false);
      (x === succ x$_1$ /\ y === succ y$_1$ /\ le$^o$ x$_1$ y$_1$ b)])
  let rec gt$^o$ x y b =
    fresh (x$_1$ y$_1$) ( conde [
      (x === zero /\ b === ^false);
      (x === succ x$_1$ /\ y === zero /\ b === ^true);
      (x === succ x$_1$ /\ y === succ y$_1$ /\ gt$^o$ x$_1$ y$_1$ b)])
  \end{lstlisting}
\end{minipage}
\end{figure*}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.7\textwidth}
\begin{lstlisting}[label={ideal:maxandlength}, caption={Ideal implementation of maxlengtho}, captionpos=b, frame=tb]
  let maxLength$^o$ xs m l = maxLength$_1^o$ xs m zero l
  let rec maxLength$_1^o$ xs m n l =
    fresh (h t l$_1$) ( conde [
      (xs === [] /\ m === n /\ l === zero);
      ((xs === h % t) /\
       (l === succ l$_1$) /\
       (conde [
         (le$^o$ h n ^true /\ maxLength$_1^o$ t m n l);
         (gt$^o$ h n ^true /\ maxLength$_1^o$ t m h l)]))])
  \end{lstlisting}
\end{minipage}
\end{figure*}

@@

\begin{table}
  \centering
  \begin{tabular}{c||c}
                   & [1..100] \\ \hline\hline
  Original         & 48.5ms  \\ \hline
  Ideal            & 32.2ms  \\ \hline
  Ideal removed    & 19.6ms  \\ \hline
  \ecce            & 23.7ms  \\ \hline
  ConsPD          & 35.7ms
  \end{tabular}

  \caption{Execution time of maxlengtho}
  \label{tbl:maxlen}
\end{table}

\includegraphics{../data/maxLengtho/max.pdf}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b, frame=tb]
  let rec eval$^o$ subst fm res = conde [
    fresh (x y z v w) (
      (fm === var v /\ elem$^o$ subst v res);
      (fm === conj x y /\ eval$^o$ st x v /\ eval$^o$ st y w /\ and$^o$ v w res);
      (fm === disj x y /\ eval$^o$ st x v /\ eval$^o$ st y w /\ or$^o$ v w res);
      (fm === neg x /\ eval$^o$ st x v /\ not$^o$ v res))]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b, frame=tb]
  let rec eval$^o$ subst fm res = conde [
    fresh (x y z v w) (
      (fm === var v /\ elem$^o$ subst v res);
      (fm === conj x y /\ and$^o$ v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
      (fm === disj x y /\ or$^o$ v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
      (fm === neg x /\ not$^o$ v res /\ eval$^o$ st x v))]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.45\textwidth}
    \begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b, frame=tb]
  let not$^o$ x y = conde [
     (x === ^true /\ y === ^false;
      x === ^false /\ y === ^true)]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.85\textwidth}
    \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b, frame=tb]
  let not$^o$ x y = nand$^o$ x x y

  let or$^o$ x y z = nand$^o$ x x xx /\ nand$^o$ y y yy /\ nand$^o$ xx yy z

  let and$^o$ x y z = nand$^o$ x y xy /\ nand$^o$ xy xy z

  let nand$^o$ a b c = conde [
    ( a === ^false /\ b === ^false /\ c === ^true );
    ( a === ^false /\ b === ^true /\ c === ^true );
    ( a === ^true /\ b === ^false /\ c === ^true );
    ( a === ^true /\ b === ^true /\ c === ^false )]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

\begin{table}
  \centering
  \begin{tabular}{c||c|c|c|c}
                   & FirstPlain & FirstNando & LastPlain & LastNando \\ \hline\hline
  Original         & 14.8ms     & 14.5ms     & 7.4ms     & 10.1ms    \\ \hline
  \ecce            & 16.0ms     & 22.3ms     & 14.0ms    & 14.9ms    \\ \hline
  ConsPD          & 9.1ms      & 9.7ms      & 9.1ms     & 9.9ms     \\
  \end{tabular}
  \caption{Execution time of evalo}
  \label{tbl:eval}
\end{table}

\includegraphics{../data/propEval/prop.pdf}

@@

% \begin{align*}
%   &type \ term = \\
%   &| \ BConst \ of \ Bool \\
%   &| \ IConst \ of \ Int \\
%   &| \ Var \ of \ Int \\
%   &| \ Plus \ of \ term * term \\
%   &| \ Mult \ of \ term * term \\
%   &| \ Eq \ of \ term * term \\
%   &| \ Lt \ of \ term * term \\
%   &| \ Let \ of \ term * term \\
%   &| \ If \ of \ term * term * term
% \end{align*}

@@

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={type:hand}, caption={A fragment of hand-written typechecker}, captionpos=b, frame=tb]
  let rec typecheck$^o$ gamma term res = conde [
    ...
    fresh (x y) (
      (term === x + y /\
       typecheck$^o$ gamma x ^(Some Integer) /\
       typecheck$^o$ gamma y ^(Some Integer) /\
       res === ^(Some Integer)));
    ...]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

\begin{figure*}[!h]
  \centering
  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={type:gen}, caption={A fragment of generated typechecker}, captionpos=b, frame=tb]
  let rec typecheck$^o$ gamma term res = conde [
    ...
    fresh (x y t1 t2) (
      (term === x + y /\
      conde [
        typecheck$^o$ gamma x ^None /\
          res === ^None;
        typecheck$^o$ gamma x ^(Some t1) /\
          typecheck$^o$ gamma y ^None /\
          res === ^None;
        typecheck$^o$ gamma x ^(Some t1) /\
          typecheck$^o$ gamma y ^(Some t2) /\
          typeEq$^o$ t1 Integer ^true /\
          typeEq$^o$ t2 Integer ^true /\
          res === ^(Some Integer);
      ])
    ...]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

>>>>

\begin{figure*}[!t]
  \centering
  \begin{minipage}{\textwidth}
    \begin{lstlisting}[label={doubleApp}, caption={Concatenation of three lists}, captionpos=b, frame=tb]
  let doubleAppend$^o$ xs ys zs res =
    fresh (ts) (append$^o$ xs ys ts /\ append$^o$ ts zs res)

  let rec append$^o$ xs ys zs = conde [
    (xs === [] /\ ys === zs);
    fresh (h t r) (xs === (h % t) /\ zs === (h % r) /\ append$^o$ t ys r)]
    \end{lstlisting}
  \end{minipage}

  \begin{minipage}{\textwidth}
    \begin{lstlisting}[label={doubleApp:ideal}, caption={Ideal implementation of concatenation of three lists}, captionpos=b, frame=tb]
      let doubleAppend$^o$ xs ys zs res = conde [
        (xs === [] /\ append$^o$ ys zs res);
        fresh (h t r) (xs === h % t /\ res === h % r /\ doubleAppend$^o$ t ys zs r)]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

\begin{figure}[!t]
  \begin{subfigure}[c]{0.35\textwidth}
    \centering
    \begin{tabular}{c||c||c}
      & forward & backward \\
      \hline\hline
      Original       & 34.8ms & 1.6ms \\ \hline
      \ecce          & 26.7ms & 1.7ms \\ \hline
      Ideal          & 26.9ms & 1.7ms \\ \hline
      ConsPD         & 47.6ms & 1.7ms
    \end{tabular}
    \caption{Evaluation results for doubleAppendo}
    \label{tbl:doubleApp}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.6\textwidth}
    \includegraphics[width=\textwidth]{../data/doubleAppendo/da700log.pdf}
    \caption{???}
    %\label{tbl:doubleApp}
  \end{subfigure}
  \caption{???}
\end{figure}

@@

@@

@@

@@

\begin{figure*}[!t]
  \centering
  \begin{minipage}{0.85\textwidth}
\begin{lstlisting}[label={cpd:maxandlength}, caption={Maximum element and length of the list}, captionpos=b, frame=tb]
  let maxLength$^o$ xs m l = max$^o$ xs m /\ length$^o$ xs l
  let rec length$^o$ xs l = fresh (h t m) ( conde [
      (xs === []    /\ l === zero);
      (xs === h % t /\ l === succ m /\  length$^o$ t m)])

@@

  let rec le$^o$ x y b = fresh (x$_1$ y$_1$) ( conde [
      (x === zero    /\ b === ^true);
      (x === succ x$_1$ /\  y === zero    /\ b === ^false);
      (x === succ x$_1$ /\  y === succ y$_1$ /\  le$^o$ x$_1$ y$_1$ b)])
  let rec gt$^o$ x y b = fresh (x$_1$ y$_1$) ( conde [
      (x === zero    /\ b === ^false);
      (x === succ x$_1$ /\  y === zero    /\ b === ^true);
      (x === succ x$_1$ /\  y === succ y$_1$ /\  gt$^o$ x$_1$ y$_1$ b)])
  \end{lstlisting}
\end{minipage}
  \begin{minipage}{0.8\textwidth}
\begin{lstlisting}[label={ideal:maxandlength}, caption={Ideal implementation of maxlengtho}, captionpos=b, frame=tb]
  let maxLength$^o$ xs m l = maxLength$_1^o$ xs m zero l
  let rec maxLength$_1^o$ xs m n l = fresh (h t l$_1$) ( conde [
      ( xs === []     /\  m === n        /\ l === zero);
      ((xs === h % t) /\ (l === succ l$_1$) /\  (conde [
         (le$^o$ h n ^true /\  maxLength$_1^o$ t m n l);
         (gt$^o$ h n ^true /\  maxLength$_1^o$ t m h l)]))])
  \end{lstlisting}
\end{minipage}
\end{figure*}

\begin{figure}[!t]
  \begin{subfigure}[c]{0.4\textwidth}
    \centering
    \begin{tabular}{c||c}
      & [1..100] \\ \hline\hline
      Original         & 48.5ms  \\ \hline
      Ideal            & 32.2ms  \\ \hline
      Ideal removed    & 19.6ms  \\ \hline
      \ecce            & 23.7ms  \\ \hline
      ConsPD          & 35.7ms
    \end{tabular}

    \caption{Execution time of maxlengtho}
    \label{tbl:maxlen}
  \end{subfigure}
  \hfill
  \begin{subfigure}[c]{0.5\textwidth}
    \includegraphics[width=\textwidth]{../data/maxLengtho/max.pdf}
    \caption{???}
    %\label{tbl:doubleApp}
  \end{subfigure}
  \caption{???}
\end{figure}

@@

@@

@@

\begin{figure*}[!t]
  \centering
  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b, frame=tb]
  let rec eval$^o$ subst fm res = conde [fresh (x y z v w) (
      (fm === var v    /\ elem$^o$ subst v res);
      (fm === conj x y /\ eval$^o$ st x v /\  eval$^o$ st y w /\  and$^o$ v w res);
      (fm === disj x y /\ eval$^o$ st x v /\  eval$^o$ st y w /\  or$^o$   v w res);
      (fm === neg x    /\ eval$^o$ st x v /\  not$^o$ v res))]
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b, frame=tb]
  let rec eval$^o$ subst fm res = conde [fresh (x y z v w) (
      (fm === var v    /\ elem$^o$ subst v res);
      (fm === conj x y /\ and$^o$ v w res /\  eval$^o$ st x v /\  eval$^o$ st y w);
      (fm === disj x y /\ or$^o$   v w res /\ eval$^o$ st x v /\  eval$^o$ st y w);
      (fm === neg x    /\ not$^o$ v res   /\ eval$^o$ st x v))]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

@@

\begin{figure*}[!t]
  \centering
  \begin{minipage}{0.7\textwidth}
    \begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b, frame=tb]
  let not$^o$ x y = conde [
     (x === ^true /\ y === ^false;
      x === ^false /\ y === ^true)]
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.9\textwidth}
    \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b, frame=tb]
  let not$^o$   x y = nand$^o$ x x y
  let or$^o$   x y z = nand$^o$ x x xx /\  nand$^o$ y y yy /\ nand$^o$ xx yy z
  let and$^o$ x y z = nand$^o$ x y xy /\   nand$^o$ xy xy z
  let nand$^o$ a b c = conde [
    ( a === ^false /\ b === ^false /\ c === ^true );
    ( a === ^false /\ b === ^true  /\ c === ^true );
    ( a === ^true  /\ b === ^false /\ c === ^true );
    ( a === ^true  /\ b === ^true  /\ c === ^false)]
    \end{lstlisting}
  \end{minipage}
\end{figure*}

@@

@@

\begin{figure}[!t]
  \centering
  \begin{subfigure}[c]{0.6\textwidth}
    \centering
    \begin{tabular}{c||c|c|c|c}
      & FirstPlain & FirstNando & LastPlain & LastNando \\ \hline\hline
      Original         & 14.8ms     & 14.5ms     & 7.4ms     & 10.1ms    \\ \hline
      \ecce            & 16.0ms     & 22.3ms     & 14.0ms    & 14.9ms    \\ \hline
      ConsPD          & 9.1ms      & 9.7ms      & 9.1ms     & 9.9ms     \\
    \end{tabular}
    \caption{Execution time of evalo}
    \label{tbl:eval}
  \end{subfigure}

  \begin{subfigure}[c]{0.8\textwidth}
    \includegraphics[width=\textwidth]{../data/propEval/prop.pdf}
    \caption{???}
  \end{subfigure}
  \caption{???}
\end{figure}

@@

@@

\begin{figure*}[!t]
  \centering
  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={type:hand}, caption={A fragment of hand-written typechecker}, captionpos=b, frame=tb]
  let rec typecheck$^o$ gamma term res = conde [
    ...
    fresh (x y) ((term === x + y /\
       typecheck$^o$ gamma x ^(Some Integer) /\
       typecheck$^o$ gamma y ^(Some Integer) /\
       res === ^(Some Integer)));
    ...]
    \end{lstlisting}
  \end{minipage}

  \begin{minipage}{0.95\textwidth}
    \begin{lstlisting}[label={type:gen}, caption={A fragment of generated typechecker}, captionpos=b, frame=tb]
  let rec typecheck$^o$ gamma term res = conde [
    ...
    fresh (x y t1 t2) ((term === x + y /\
      conde [
        typecheck$^o$ gamma x ^None       /\ res === ^None;
        typecheck$^o$ gamma x ^(Some t1) /\
          typecheck$^o$ gamma y ^None     /\ res === ^None;
        typecheck$^o$ gamma x ^(Some t1) /\  typecheck$^o$ gamma y ^(Some t2) /\
          typeEq$^o$ t1 Integer ^true    /\ typeEq$^o$ t2 Integer ^true /\
          res === ^(Some Integer);
      ])
    ...]
    \end{lstlisting}
  \end{minipage}
\end{figure*}
